The \dealii{} Library, Version 9.3

### 2021-03-07 9.3/paper.tex
Update simplex section

@@

\begin{table}[!h]
  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
  reference-cell types. Vectorial elements can be constructed based on these classes
  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
  \centering
  \begin{tabular}{cccc}
    \toprule
    \textbf{reference cell} & \textbf{finite element} & \textbf{dimension} & \textbf{degree}\\
    \midrule
    simplex (line, triangle, tetrahedron) & FE\_P, FE\_DGP  & 1--3 & 1--2 \\
    pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
    wedge & FE\_WedgeP, FE\_WedgeDGP  & 3 & 1--2 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The internal data structures of \texttt{DoFHandler} used to be hard-coded for pure
hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
CRS-like data structures. Due to the need of CRS data structures in
\texttt{DoFHandler} in the context of more general meshes, we have merged
\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons. For
more details see Subsection~\ref{TODO}.

@@

\begin{c++}
Simplex::FE_P<dim, spacedim> fe(degree);
MappingFE<dim, spacedim> mapping(Simplex::FE_P<dim, spacedim>(1));
Simplex::QGauss<dim> quad(degree + 1);

@@

FEValues<dim, spacedim> fe_values(mapping, fe, quad, flags);
\end{c++}
The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
only linear iso-parametric mapping (\texttt{MappingFE}) is available, which takes one of the 
finite elements mentioned before. For quadrature, the classes \texttt{QGauss}, 
\texttt{QGaussPyamid}, and \texttt{QGaussWedge} from the \texttt{simplex} namespace are
available.

\subsubsection{Mixed mesh}
For mixed meshes, concepts known from the hp context have been applied: 
 different finite-element classes are assigned to different cell types. In the 
case of a 2D mixed mesh, which can only consists of triangles and
quadrilaterals, the finite element defined on a triangle (e.g., \texttt{Simplex::FE\_P})
and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:

\begin{c++}  
hp::FECollection<dim, spacedim> fe
  {Simplex::FE_P<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
\end{c++}

@@

The introduction of simplex and mixed meshes leads to some implications
for the user if these features should be used. For instance, each cell might have a different type with 
different number of vertices, lines, and faces so that these quantities cannot be 
compile-time constants anymore. This information used to be queried from
the \texttt{GeometryInfo} class. To avoid using this class, we have extended
relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
\texttt{n\_faces()} so that users can simply write:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(unsigned int f = 0; f < cell->n_faces(); ++f)
    (void) cell->face(f);
\end{c++}
Alternatively, one can use an iterator-based approach to loop over all faces
of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
various number of faces. 
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(const auto & face : cell->face_iterators())
    (void) face;
\end{c++}
Furthermore, the number of degrees of freedom might differ between cells so that
local data storage units needed for assembly might need to be resized for each 
cell as it is already good practice in the hp context:
\begin{c++}
std::vector<double> local_rhs;
for(const auto & cell : tria.active_cell_iterators())
 {
   hp_fe_values.reinit(cell);
   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
 }
\end{c++}

@@

We tried---where possible---to provide utility functions so that users do not need to access a \texttt{GeometryInfo}-like data structure. However, these functions internally rely on the \texttt{ReferenceCell::\allowbreak internal::\allowbreak Info} classes, which are similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells. Users can query the reference-cell type of geometric entity via:
\begin{c++}
const auto reference_cell_type = cell->reference_cell_type();
\end{c++}
and can access the relevant struct describing the current reference cell or its i-th face via:
\begin{c++}
ReferenceCell::internal::Info::get_cell(reference_cell_type);
ReferenceCell::internal::Info::get_face(reference_cell_type, i);
\end{c++}

@@

@@

Until now, \dealii has only supported local smoothing multigrid algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}. With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}) and discontinuous (\texttt{FE\_DGQ}) elements  in a geometric and a polynomial sense has been added. These multigrid variants promise less iteration numbers and better parallel scalability with the disadvantage of having to deal with hanging nodes during smoothing and of more work per iteration overall.

@@

Internally, the transfer operators categorize fine/coarse cell pairs according to the 
occurrence of refinement 
(refinement or no-refinement) and to the polynomial-degree pair ($k_f$-$k_c$) into groups. 
The projection and restriction are performed with efficient matrix-free kernels applying
1D projection and restriction matrices.

@@

The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
templated to reach high performance. In particular, the template parameters include
the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
For applications codes that rely on operators of many different degrees (e.g., because
they use p-multigrid or hp algorithms), the instantiation
means a long compilation time.

In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations, which do
not rely on the template parameters $k$ and $q$ (expressed in the code
with "-1" and "0"):
\begin{c++}
FEEvaluation<dim, -1, 0, n_components, Number, VectorizedArrayType>
  phi(range, dofhandler_index, quadrature_index, first_selected_component);
\end{c++}
The meaning of the new first argument is 
discussed in Subsubsection~\ref{subsubsection:mf:hp}.

@@

The novel tutorial \texttt{step-67b} shows the usage of the new MPI-3.0 feature
 in the context of the solution
of the Euler equations. Here, a speed-up of 27\% could be reached compared to the
regular version \texttt{step-67}, by using the new feature.
For more details and a successful application in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.

>>>>

\usepackage{multirow}

@@

\begin{table}[!h]
  \caption{\it List of new scalar \texttt{FiniteElement} classes for the new
  reference-cell types. Vectorial elements can be constructed based on these classes
  via \texttt{FE\_Systems}.}\label{tab:simplex:fe}
  \centering
  \begin{tabular}{cccc}
    \toprule
    \textbf{reference cell} & \textbf{finite element} & \textbf{dimension} & \textbf{degree}\\
    \midrule
    \multirow{2}{*}{simplex (line, triangle, tetrahedron)} & FE\_SimplexP, FE\_SimplexDGP  & 1--3 & 1--2 \\
                                          & FE\_SimplexP\_Bubbles  & 1--3 & 1--2 \\
    pyramid & FE\_PyramidP, FE\_PyramidDGP  & 3 & 1 \\
    wedge & FE\_WedgeP, FE\_WedgeDGP  & 3 & 1--2 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

The internal data structures of \texttt{DoFHandler} used to be hard-coded for pure
hypercube meshes, while the \texttt{hp::DoFHandler} used to be built around
CRS-like data structures. Due to the need of CRS data structures in the
\texttt{DoFHandler} in the context of more general meshes, we have merged
\texttt{hp::DoFHandler} into the \texttt{DoFHandler}. The class \texttt{hp::DoFHandler}, which is now a dummy derivation of \texttt{DoFHandler}, currently only exists for compatibility reasons. For
more details see Subsection~\ref{TODO}.

@@

\begin{c++}
FE_SimplexP<dim, spacedim> fe(degree);
MappingFE<dim, spacedim> mapping(FE_SimplexP<dim, spacedim>(1));
QGaussSimplex<dim> quad(degree + 1);

@@

FEValues<dim, spacedim> fe_values(mapping, fe, quad, flags);
\end{c++}
The  list of currently supported finite-element classes is provided in Table~\ref{tab:simplex:fe}. Currently,
only linear iso-parametric mapping (\texttt{MappingFE} and \texttt{MappingFEFields}) is available. For quadrature, the classes \texttt{QGaussSimplex}, \texttt{QWitherdenVincentSimplex},  
\texttt{QGaussPyamid}, and \texttt{QGaussWedge} are
available.

\subsubsection{Mixed mesh}
For mixed meshes, concepts known from the hp context have been applied: 
 different finite-element classes are assigned to different cell types. In the 
case of a 2D mixed mesh, which can only consists of triangles and
quadrilaterals, the finite element defined on a triangle (e.g., \texttt{FE\_SimplexP})
and on a quadrilateral (e.g., \texttt{FE\_Q}) can be collected in a \texttt{hp::FECollection}:

\begin{c++}  
hp::FECollection<dim, spacedim> fe
  {FE_SimplexP<dim, spacedim>(degree), FE_Q<dim, spacedim>(degree)};
\end{c++}

@@

The introduction of simplex and mixed meshes leads to some implications
for the user if these features should be used. For instance, each cell might have a different type with 
different number of vertices, lines, and faces so that these quantities can not be 
compile-time constants anymore. This information used to be queried from
the \texttt{GeometryInfo} class. To avoid using this class, we have extended
relevant classes, e.g., \texttt{TriaAccessor} or \texttt{TriaCellAccessor},
with useful new functions like \texttt{n\_vertices()}, \texttt{n\_lines()}, or
\texttt{n\_faces()} so that users can simply write:
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(unsigned int f = 0; f < cell->n_faces(); ++f)
    (void) cell->face(f);
\end{c++}
Alternatively, one can use an iterator-based approach to loop over all faces
of a cell, introduced in the last release. The relevant functions have been adjusted to be able to deal with
various number of faces. 
\begin{c++}
for(const auto & cell : tria.active_cell_iterators())
  for(const auto & face : cell->face_iterators())
    (void) face;
\end{c++}
Furthermore, the number of degrees of freedom might differ between cells so that
local data storage units needed for assembly might need to be resized for each 
cell as it is already good practice in the hp context:
\begin{c++}
std::vector<double> local_rhs;
for(const auto & cell : tria.active_cell_iterators())
 {
   hp_fe_values.reinit(cell);
   local_rhs.resize(hp_fe_values.get_present_fe_values().n_dofs_per_cell());
 }
\end{c++}

@@

We tried---where possible---to provide utility functions so that users do not need to access a \texttt{GeometryInfo}-like data structure. However, these functions internally rely on the \texttt{ReferenceCell} class, which is similarly structured as \texttt{GeometryInfo} and even use it for hypercube cells. Users can query the right \texttt{ReferenceCell} of a geometric entity via:
\begin{c++}
const auto reference_cell = cell->reference_cell();
\end{c++}
and of its i-th face via:
\begin{c++}
const auto face_reference_cell = reference_cell.face_reference_cell(i);
\end{c++}

@@

We have created two new tutorials presenting the new simplex (\texttt{step-3b}) and mixed mesh features
(\texttt{step-3c}). Similarly as \texttt{step-3}, they solve a Poisson problem, however, focus
on the changed workflow.

By the time of writing, there have been 74 tests (in the folder \texttt{tests/simplex}) 
targeting the new simplex and mixed mesh support. In particular, the folder also
contains ported variants of the following tutorials: 1, 2, 3, 4, 6, 7, 8, 12, 17, 18, 20, 23, 31, 38,
40, 55, 67, 68, and 74. These might be also good starting points.

@@

Until now, \dealii has only supported local smoothing multigrid algorithms~\citep{ClevengerHeisterKanschatKronbichler2019}. With release 9.3, the support for global coarsening of continuous (\texttt{FE\_Q}, \texttt{FE\_SimplexP}) and discontinuous (\texttt{FE\_DGQ}, \texttt{FE\_SimplexDGP}) elements  in a geometric and a polynomial sense has been added. These multigrid variants promise less iteration numbers and better parallel scalability with the disadvantage of having to deal with hanging nodes during smoothing and of more work per iteration overall.

@@

Internally, the transfer operators categorize fine/coarse cell pairs according to the 
occurrence of refinement 
(refinement or no-refinement) and to the polynomial-degree pair ($k_f$-$k_c$) into groups. 
The projection and restriction are performed with efficient matrix-free kernels applying
1D projection and restriction matrices in the case of hypercube cells and via full projection and
restriction matrices in the case of simplex cells.

@@

The classes \texttt{FEEvaluation} and \texttt{FEFaceEvaluation} are highly
templated to reach high performance. In particular, the template parameters include
the polynomial degree of the finite element $k$ and the number of the 1D quadrature points $q$.
For application codes that rely on operators of many different degrees (e.g., because
they use p-multigrid or hp algorithms), the instantiation
means a long compilation time.

In the current release, we have improved the \texttt{FEEvaluation} and the \texttt{FEFaceEvaluation} class implementations that do
not rely on the template parameters $k$ and $q$ (expressed in the code
with "-1" and "0"):
\begin{c++}
FEEvaluation<dim, -1, 0, n_components, Number, VectorizedArrayType>
  phi(range, dofhandler_index, quadrature_index, first_selected_component);
\end{c++}
The meaning of the new first argument is 
discussed in Subsubsection~\ref{subsubsection:mf:hp}.

@@

The novel tutorial \texttt{step-76} shows the usage of the new MPI-3.0 feature
 in the context of the solution
of the Euler equations. Here, a speed-up of 27\% could be reached compared to the
original version, \texttt{step-67}, by using the new feature.
For more details and a successful application in the library \texttt{hyper.deal}, see \citep{munch2020hyperdeal}.
